\section{Day 10: Third Application of Cauchy's Formula; Analytic Extension of Gamma Function (Oct. 2, 2025)}
We now discuss a third application of the Cauchy integral formula. Let $f$ be given by
\[ f(z) = \int_a^b F(z, s) \, ds, \]
where $F(z, s)$ is holomorphic for $z \in \CC$ and continuous in $s \in \RR$. Is $f$ holomorphic in $z$?
\begin{example}
    Consider the gamma function $\Gamma(s) = \int_0^\infty e^{-t} t^{s-1} \, dt$, which can be thought of as
    \[ \lim_{\eps \to 0^+} \int_\eps^{\frac{1}{\eps}} e^{-t} t^{s-t} \, dt. \]
    For $s > 0$, we have convergence, as for $t \to 0$, we have that $e^{-t} t^{s-1} \approx t^{s-1}$, which is integrable; for large $t$, the convergence is essentially ``guaranteed'' by the exponential decay of the integrand, i.e., $\abs{e^{-t} t^{s-1}} \leq C e^{-\frac{t}{2}}$, which is integrable over $\eps\inv$ to infinity. We ask, is $\Gamma(s)$ convergent or analytic when $\Re s > 0$?
\end{example}
\begin{theorem}[Thm.\ 5.4, Shakarchi]
    Let $F(z, s)$ be defined for $(z, s) \in \Omega \times [0, 1]$, where $\Omega$ is open in $\CC$. Suppose $F$ satisfies the following,
    \begin{enumerate}[(i)]
        \item $F(z, s)$ is holomorphic in $z$ for all $s \in [0, 1]$,
        \item $F$ is continuous on $\Omega \times [0, 1]$.
    \end{enumerate}
    Then $f(z) = \int_0^1 F(z, s) \, ds$ is holomorphic on $\Omega$.
\end{theorem}
\begin{proof}
    Observe that we may write $\int_0^1 F(z, s) \, ds$ in terms of Riemann sums. Define $\{f_n\}_{n \in \NN}$ on $\Omega$ by
    \[ f_n(z) = \frac{1}{n} \sum_{k=1}^n F(z, k/n), \]
    for which we may note that each $f_n$ is holomorphic in $z$ by property (i). We claim that $\{f_n\}$ converges uniformly to some $f$, i.e., as $n \to \infty$, we have that
    \[ \frac{1}{n} \sum_{k=1}^n F(z, k/n) \to \int_0^1 F(z, s) \, ds = f(z). \]
    For any open disc $D$ with closure $\ol D \subset \Omega$, we want to show that $\{f_n\} \to f$ uniformly on $D$. Since $F$ is continuous on $\Omega \times [0, 1]$, it is also continuous on $\ol D \times [0, 1]$, which we observe is compact, so we indeed have uniform continuity. In particular, this means that for all $\eps > 0$, there exists $\delta > 0$ such that $\sup_{z \in D} \abs{F(z, s_1) - F(z, s_2)} < \eps$ whenever $\abs{s_1 - s_2} < \delta$.
    \\[8pt]
    In this manner, we may take that given $n > \delta\inv$ and $z \in D$, we have that
    \begin{align*}
        \abs{f_n(z) - f(z)} &= \abs{\sum_{k=1}^n \int_{(k-1)/n}^{k/n} F(z, k/n) - F(z, s) \, ds} \\
        &\leq \sum_{k=1}^n \int_{(k-1)/n}^{k/n} \abs{F(z, k/n) - F(z, s)} \, ds \tag{Triangle ineq.} \\
        &< \sum_{k=1}^n \int_{(k-1)/n}^{k/n} \eps \, ds = \sum_{k=1}^n \frac{\eps}{n} = \eps.
    \end{align*}
    Thus, we establish that $f_n \to f$ on $D$, and by the first consequence of Cauchy's integral formula (equivalently, Theorem 5.2 in Shakarchi and 9.3 in these notes), we have that $f$ is holomorphic in $z$ on $\Omega \times [0, 1]$.
\end{proof}
\noindent We now return to the Gamma function, which will follow Ch.\ 6 in Shakarchi.
\begin{proposition}[Prop.\ 1.1, Shakarchi; p.\ 160]
    $\Gamma(s)$ extends analytically on $\Re s > 0$, and is given by the same formula.
\end{proposition}
\begin{proof}
    It suffices to consider that this holds for every strip $S_{\delta, m} = \{s \in \CC \mid \delta < \Re s < M\}$, where $\delta, M$ are positive reals satisfying $\delta < M$. For all $s \in S_{\delta, M}$, denote $\Re s$ by $\sigma$. Then
    \[ \abs{\int_0^\infty e^{-t} t^{s-1} \, dt} \leq \int_0^\infty \abs{e^{-t} t^{s-1}} \, dt = \int_0^\infty e^{-t} t^{\sigma - 1} \, dt, \]
    per Euler's formula, of which we may observe the latter integral converges by real convergence (as discussed earlier). Now, let $f(s, t) = e^{-t} t^{s-1}$ be holomorphic in $s$ and continuous in $(s, t)$. For $\eps > 0$, define
    \[ F_\eps(s) = \int_\eps^{\frac{1}{\eps}} e^{-t} t^{s-1} \, dt; \]
    per our previous result, we have that $F_\eps(s)$ is holomorphic in $s$ on $S_{\delta, M}$. Also, per the first result of Cauchy's formula, it suffices to check that $F_\eps$ converges uniformly to $\Gamma$ on $S_{\delta, M}$. Directly write as follows,
    \[ \abs{\Gamma(s) - F_\eps(s)} \leq \int_0^\eps e^{-t} t^{\sigma - 1} \, dt + \int_{\frac{1}{\eps}}^\infty e^{-t} t^{\sigma - 1} \, dt, \]
    for which we may observe that
    \begin{align*}
        \int_0^\eps e^{-t} t^{\sigma - 1} \, dt &= \frac{\eps^\delta}{\delta} \taking{\eps \to 0} 0 \tag{on $0 < \eps < 1$}, \\
        \abs{\int_{\frac{1}{\eps}}^\infty e^{-t} t^{\sigma - 1} \, dt} &\leq \int_{\frac{1}{\eps}}^\infty e^{-t} t^{M-1} \, dt \leq C \int_{\frac{1}{\eps}}^\infty e^{-\frac{t}{2}} \, dt \taking{\eps \to 0} 0.
    \end{align*}
    Thus, we obtain uniform convergence of $F_\eps \to \Gamma$ independent of $s$, and we indeed have an analytic continuation of $\Gamma$ to the half-plane $\Re s > 0$.
\end{proof}
\noindent Let $\zeta(s) = \sum_{n=1}^\infty n^{-s}$ be the Riemann $\zeta$ function, defined on $\Re s > 1$ (clearly, otherwise the series does not converge). $\zeta$ is analytic in the half-plane $\Re s > 1$, and we may show that it has an analytic extension using $\Gamma$. We know $\Gamma$ can be analytically extended to a meromorphic function on $\CC$ with simple poles at $s = 0, -1, -2, \dots$ using the relation $\Gamma(s + 1) = s \Gamma(s)$ and the fact that $\Gamma(s)\inv$ is entire. Then we have that
\[ \eta(s) = \pi^{-\frac{s}{2}} \Gamma(s/2) \zeta(s). \]