\section{Day 10: }
Last class, we defined submanifolds, the rank of a smooth map, and maps of maximum rank (submersions) $f \colon M^m \to N^n$, where $M$ is of dimension $m$ and $N$ is of dimension $n$ with $m \geq n$.
\begin{fact}
    In a submersion $f$, $f\inv(p)$ are submanifolds.
\end{fact}
Recall that, if $f \colon M^m \to N^n$ with $m \leq n$ has rank $m$ everywhere, then it is an \textit{immersion}.
\\[-12pt]
\begin{theorem}[Normal form theorem]
    Let $f \colon M \to N$ be a map of rank $m$ at $p \in M$. Then there are charts $(U, \varphi)$ at $p \in M$ and $(V, \psi)$ at $f(p)$ for $N$ such that the diagram commutes,
    \[ \begin{tikzcd} U \arrow[r, "f"] \arrow[d, "\varphi"] & V \arrow[d, "\psi"] \\ \varphi(U) \arrow[r, hook] & \psi(V) \end{tikzcd} \]
\end{theorem}
\begin{proof}
    This follows from the $\RR^n$ version.
\end{proof}
Equivalently, $\restr{f}{U}$ is the inclusion of a submanifold, which we see by restricting $(V, \psi)$ to $\psi\inv(U \times \RR^{n-m})$. We now discuss examples.
\begin{enumerate}[(i)]
    \item A curve is a smooth map $\gamma \colon (a, b) \to M$. A curve is immmersed at $p$ if, for $(V, \psi)$ a chart, $\frac{d}{dt} \restr{(\psi \circ \gamma)}{t} \neq 0$. However, if we consider $\gamma(t) = (t^2, t^3)$ or $(t^3, t^3)$ in $\RR^2$, the image \textit{is} a submanifold, but the parameterization is bad!
    \item Consider the figure eight, given by the map $S^1 \to \RR^2$.
    \item Consider $\RR \to \RR^2$ given by $t \mapsto e^{it}$.
    \item (\S4.42) something something immersion to get a klein bottle.
\end{enumerate}
\begin{definition}
    An \textit{embedding} is an injective immersion, where the image is a submanifold.
\end{definition}
\begin{theorem}
    If $M$ is compact, then any injective immersion $f \colon M \to N$ is an embedding.
\end{theorem}
\begin{proof}
    We want to show that, at each point $p \in M$, there is a submanifold chart $(V, \psi)$, i.e., 
    \[ \begin{tikzcd} V \cap f(M) \arrow[r, hook] \arrow[d, "\restr{\psi}{M}"] & V \arrow[d, "\psi"] \\ \psi(V) \cap \RR^m \arrow[r] & \psi(V) \end{tikzcd} \]
    To start, we know that there exists charts $(U, \varphi)$ and $(V, \psi)$ such that
    \[ \begin{tikzcd} U \arrow[r, "f"] \arrow[d, "\varphi"] & V \arrow[d, "\psi"] \\ \psi(V) \cap \RR^m \arrow[r] & \psi(V) \end{tikzcd} \]
    Recall that the images of compact sets are compact. This means $f(M)$ is compact in $N$, whence $f(M \setminus U) = A$ is closed (because closed subsets of compact sets are compact). Thus, $V \setminus A$ is open, and so $f(U) = (V \setminus A) \cap f(M)$. From this, we get the diagram
    \[ \begin{tikzcd} (V \setminus A) \cap f(M) \arrow[r] \arrow[d, "\restr{\psi}{f(M)}"] & V \setminus A \arrow[d, "\psi"] \\ \psi(V) \cap \RR^m \arrow[r] & \psi(V) \end{tikzcd} \qedhere \]
\end{proof}
We give an example. Consider $A \neq B \neq C \in \RR$; then define
\[ F_{A,B,C} \colon \RP^2 \to \RR^4, \quad \pcoor{\alpha : \beta : \gamma} \mapsto (\beta \gamma, \alpha \gamma, \alpha \beta, A\alpha^2, B\beta^2, C\gamma^2), \]
where $(\alpha, \beta, \gamma) \in S^2$, is an injective immersion, i.e., an embedding, since $\RP^2$ is compact.